\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-23 appendices.tex
Adding a bib file

%% \documentclass[11pt]{article}
\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

%LatexiDiff ignore tikz:
% $: latexdiff -c "PICTUREENV=(?:picture|tikzpicture|DIFnomarkup)[\w\d*@]*" old.tex new.tex > diff.tex

%\IEEEoverridecommandlockouts                              %

%\usepackage{mathcomSTEP}

%\overrideIEEEmargins 
%\IEEEoverridecommandlockouts                              % This command is only needed if 
% you want to use the \thanks command

%\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\onecolumn

\usepackage{mathptmx} 
\usepackage{times} 
\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}

%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

\usepackage[hidelinks]{hyperref}
%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
%\usepackage{caption}
\usepackage{epstopdf}

\newtheorem{thm}{\bf{Theorem}}
\newtheorem{cor}[thm]{\bf {Corollary}}
\newtheorem{lem}[thm]{\bf {Lemma}}
\newtheorem{prop}[thm]{\bf {Proposition}}
\newtheorem{example}{\bf {Example}}
\newtheorem{definition}{\bf {Definition}}
\newtheorem{rem}{\bf {Remark}}
\newtheorem{assump}{\bf {Assumptions}}

\newcommand{\ex}[1]{\ensuremath{\mathbb{E}\left[ #1\right]}}

\newcommand{\Var}[1]{\ensuremath{\mathrm{Var} \left[ #1\right]}}

\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\Prob}{\mathbb{P} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}
\newcommand{\CEO}{\mathsf{CEO}}
\newcommand{\Ncal}{\mathcal{N}}

\newcommand{\enc}{\mathsf{enc}}
\newcommand{\est}{\mathsf{est}}

\tikzstyle{int}=[draw, fill=blue!10, minimum height = 1cm, minimum width=1.5cm,thick ]
\tikzstyle{sint}=[draw, fill=blue!10, minimum height = 0.5cm, minimum width=0.8cm,thick ]
\tikzstyle{sum}=[circle, fill=blue!10, draw=black,line width=1pt,minimum size = 0.5cm, thick ]
\tikzstyle{ssum}=[circle, fill=blue!10,draw=black,line width=1pt,minimum size = 0.1cm]
\tikzstyle{int1}=[draw, fill=blue!10, minimum height = 0.5cm, minimum width=0.7cm,thick ]
\tikzstyle{enc}=[draw, fill=blue!10, minimum height = 2.7cm, minimum width=1cm,thick ]
\tikzstyle{int}=[draw, fill=blue!10, minimum height = 1cm, minimum width=1.5cm,thick ]

\author{
    \IEEEauthorblockN{
Alon Kipnis\IEEEauthorrefmark{1} and   
    John C. Duchi\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}  \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Stanford University, Department of Statistics \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Stanford University, Department of Electrical Engineering. }
}

\title{\LARGE \bf Mean Estimation from One-Bit Measurements}

\begin{document}
\graphicspath{{./Figs/}}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{abstract}

%{\color{red}  See LASSO results in http://arxiv.org/abs/1506.02181v1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{introduction}

\section{Problem Formulation \label{sec:problem}}

Let $f(x)$ be a symmetric and log-concave density function with a finite second moment $\sigma^2$. For $\theta \in \Theta$, denote by $P_X$ the probability distribution with density $f\left( x-\theta \right)$. Therefore, $P_X$ is an absolutely continuous log-concave distribution with mean $\theta$ and variance $\sigma^2$. Symmetry and log-concavity of $f(x)$ imply that $P_X$ is strongly unimodal (has a unique local maxima) with its mode at $x =\theta$ \cite{ibragimov1956composition}. We further assume that the \emph{parameter space} $\Theta$ is a closed interval of the real line. 
\par
In some situations it is useful to assume that $\theta$ is drawn once from the prior distribution $\pi$ on $\Theta$. In this case we assume that $\pi$ is an absolutely continuous distribution, and denote its density by $\pi(\theta)$, i.e., $\pi(d\theta) = \pi(\theta)d\theta$.  
\par
The random variables $X_1,\ldots,X_n$ represent $n$ independent samples from $P_X$. 
We are interested in estimating $\theta$ from a set of $n$ binary messages $B_1,\ldots,B_n$, obtained from $X_1,\ldots,X_n$ under three possible scenarios illustrated in Figure~\ref{fig:setup}: 
\begin{itemize}
\item[(i)~~] Centralized $B_i(X_1,\ldots,X_n)$, $i=1,\ldots,n$.
\item[(ii)~] Adaptive $B_i(X_i,B_1,\ldots,B_{i-1})$, $i=2,\ldots,n$.
\item[(iii)] Distributed $B_i(X_i)$, $i=1,\ldots,n$.
\end{itemize}

The performance of an estimator ${\theta}_n \triangleq {\theta}_n(B^n)$ in any of these cases is measured according to the quadratic risk:
\begin{equation}
\label{eq:error_def}
R_n \triangleq \mathbb E\left({\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X_1,\ldots,X_n$ and, whenever available, a prior distribution $\pi(\theta)$ over $\Theta$.  
%It is well known that minmax estimation error 
%\begin{equation}
%R_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left({\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $R_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}. 
%
The main problems we consider in this paper are the minimal value of \eqref{eq:error_def}, as a function of $n$ and $f(x)$, under different choices of the encoding functions in cases (i), (ii), and (iii). \par

We give particular attention to the ARE of estimators with respect to an asymptotically normal efficient estimator that is not subject to the bit constraint. Specifically, let $\{a_n,\,n\in \mathbb N\}$ be a sequence such that 
\[
\sqrt{a_n}\left({\theta}_n - \theta\right) \overset{d}{\longrightarrow} \Ncal(0, \sigma^2).
\]
Then the ARE of ${\theta}_n$ with respect to an unconstrained efficient estimator for $\theta$ is defined as \cite[Def. 6.6.6]{lehmann2006theory}
\[
\ARE({\theta}_n) \triangleq
\lim_{n\rightarrow \infty} \frac{a_n}{n}. 
\]
Note that in the special case where there exists $V \in \mathbb R$ such that
\[
a_n \mathbb E \left({\theta}_n - \theta \right)^2 = V + o(1),
\]
with $o(1)\to 0$ as $n\to \infty$, then the ARE of ${\theta}_n$ is finite and equals $\sigma^2/V$. \par

In addition to the notation above, we also denote by $F(x)$ the cumulative distribution function of $X_i$, and define
\begin{equation} \label{eq:eta_def}
\eta(x) \triangleq \frac{f^2(x)}{F(x)(1-F(x))} =  \frac{f(x)f(-x)}{F(x)F(-x)}, 
\end{equation}
where the last equality is due to symmetry of $f(x)$. We note that 
\begin{equation}
\label{eq:eta_h}
\eta(x) = h(x)h(-x) = f(x) \left( h(x) + h(-x) \right), 
\end{equation}
where 
\[
h(x) \triangleq \frac{f(x)}{1-F(x)} = \frac{f(x)}{F(-x)}
\]
is the \emph{hazard} function (a.k.a. \emph{failure rate} or \emph{force of mortality}), which is a monotone increasing function since $f(x)$ is log-concave \cite{bagnoli2005log}. 
%
For $f(x)$ the normal density, it is shown in \cite{Samford1953} and \cite{hammersley1950estimating} that $\eta(x)$ is a strictly decreasing function of $|x|$, as illustrated in Fig.~\ref{fig:eta}. 
%
In this paper we only consider the normal distribution and other log-concave symmetric distributions for which this property of $\eta(x)$ holds. Specifically, we require the following:
\begin{assump} \label{assump:failure_rate}
 $\eta(x)$ has a unique maximum (at the origin) and is a non-increasing function of $|x|$. 
\end{assump}
%That is, Assumption~\ref{assump:failure_rate} says that $\eta'(x) < 0 $ for all $x>0$. 
Under this assumption we have, 
\[
4f^2(x) \leq \eta(x) \leq \eta(0),
\] 
%
where $\eta(0) = 4 f^2(0)$ is the asymptotic variance of the sample median. Combined with log-concavity of $f(x)$, Assumption~\ref{assump:failure_rate} implies that $\eta(x)$ vanishes as $|x|\rightarrow \infty$. \par
%
Assumption~\ref{assump:failure_rate} is satisfied, for example, by the generalized normal distributions with a shape parameter between $1$ and $2$ (including normal and Laplace distributions). Symmetric log-concave distributions that do not satisfy Assumption~\ref{assump:failure_rate} include the uniform distribution and the generalized normal distribution with shape parameter greater than $2$.\par

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=8cm, height=6cm,
xmin = -3, xmax=3, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 0.9,
samples=10, 
xlabel= $x$,
xtick={-2,-1,0,1,2},
xticklabels={-2,-1,0,1,2},
ytick={0,0.3989423,0.6366198},
yticklabels={0,$\frac{1}{\sqrt{2 \pi}}$,$2/\pi$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, solid, smooth] plot table [x = x, y = y, col sep=comma] {./Figs/eta.csv};
\addlegendentry{$\eta(x)$};
\addplot[domain = -5:5, samples = 50, color = red, solid, smooth]  {exp(-x^2/2) / sqrt(2*3.14159)};
\addlegendentry{$\phi(x)$};
\addplot[domain = -5:5, samples = 50, color = black, solid, dashed]  {4*exp(-x^2) / (2*3.14159)};
\addlegendentry{$4\phi^2(x)$};

\end{axis}
\end{tikzpicture}
%
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=8cm, height=6cm,
xmin = -4, xmax=4, 
restrict y to domain = -10:0,
ymin = -10,
samples=10, 
xlabel= $x$,
xtick={-3,-2,-1,0,1,2,3},
xticklabels={-3,-2,-1,0,1,2,3},
ytick={0,-0.45158,-0.919},
yticklabels={0,,},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, solid, smooth] plot table [x = x, y = logy, col sep=comma] {./Figs/eta.csv};
\addlegendentry{$\log \eta(x)$};
\addplot[domain = -5:5, samples = 30, color = red, solid, smooth]  {-(x)^2/2 -0.9189};
\addlegendentry{$\log \phi(x)$};
\end{axis}
\end{tikzpicture}
\caption{
The function $\eta(x) = f^2(x) / F(x)F(-x)$ for $f(x) = \phi(x)$ the standard normal density.
\label{fig:eta}
}
\end{center}
\end{figure}

\section{Consistent Estimation and Off-the-shelf Bounds \label{sec:preliminary}}
%
In this section we adapt known results to derive lower and upper bounds on the relative efficiency under setting (iii). These bounds establish the following facts:
\begin{itemize}
\item[I.] A consistent estimator with an asymptotically normal distribution always exists in setting (iii), and hence in setting (ii). 
\item[II.] For the normal distribution, the ARE in setting (iii) is at most $3/4$. Namely, under setting (iii), all estimators are strictly inferior compared to the sample mean. 
\end{itemize}

\subsection{Consistent Estimation}
Fix $\theta_0 \in \mathbb R$ and define the $i$th message by 
\[
B_i = \mathbf 1_{X_i>\theta_0}, 
\]
where $\mathbf 1_A$ is the indicator of the event $A$. We have
\[
P_n \triangleq\frac{1}{n} \sum_{i=1}^n B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),  
\]
so that 
\begin{equation}
\label{eq:estimator_naive}
{\theta}_n = \theta_0 + F^{-1}\left( P_n \right)
\end{equation}
is a consistent estimator for $\theta$ in the distributed setting of Figure~\ref{fig:setup}-(iii), where we note that $F(x)$ is invertible over the support of $f(x)$ which is a connected set due to log-concavity. Furthermore, the variance of $P_n$ is $F(\theta-\theta_0)\left(1-F(\theta-\theta_0)\right)$, and hence the delta method implies that ${\theta}_n$ is asymptotically normal with variance 
\begin{equation} 
 \frac{F(\theta-\theta_0)\left(1-F(\theta-\theta_0)\right)}{f^2(\theta-\theta_0)} = \frac{1}{
\eta(\theta-\theta_0)}.
\end{equation}
In particular, the ARE of ${\theta}_n$ equals $\eta(\theta - \theta_0)\sigma^2$. In other words, for a prescribed accuracy, ${\theta}_n$ of  \eqref{eq:estimator_naive} estimates $\theta$ with sample size that is $\eta(\theta- \theta_0)  \sigma^2$ times the samples size required for the sample mean. \par
%
%
Assumption~\ref{assump:failure_rate} implies that for  all $n$ large enough the ARE of ${\theta}_n$ of \eqref{eq:estimator_naive} is never greater than $\eta(0)$, which is the ARE of the sample median. This ARE is attained only when $\theta_0 = \theta$, whereas $\theta$ is apriori unknown. Since $\eta(x)$ vanishes as $|x|\rightarrow \infty$, the ARE of ${\theta}_n$ may be very small when $\theta$ is away from $\theta_0$. As an example, when $f(x)$ is a normal density, the ARE of ${\theta}_n$ is less than $\approx 0.14$ when $\theta_0$ is $2$ standard deviations from $\theta$. Therefore, the estimator ${\theta}_n$ has little practical value unless the radius of $\Theta$ is small compared to the standard deviation. \par 
It is suggested that we can attain lower variance in estimation by updating the threshold value $\theta_0$ in \eqref{eq:estimator_naive} after observing one or a batch of the single bit messages. Schemes with such adaptation fall within the adaptive setting of Figure~\ref{fig:setup}-(ii) which we consider in Section~\ref{sec:sequential}. \par

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k$ goes to infinity were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for quadratic risk attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

\begin{proof}
See Appendix~\ref{app:proof:prop_ceo_lower_bound}.
\end{proof}

From the formulation of the CEO problem, it follows that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed setting (case (iii)) is attributed to the ability to perform coding over blocks. Namely, each encoder in the CEO setting may encode an arbitrary number of $k$ independent realizations of $\theta$ using $k$ bits, versus only one realization with one bit in ours. In other words, it is the ability to exploit the geometry of a high-dimensional product probability space that distinguishes between the CEO problem with one bit per encoder on average and the mean estimation problem from one-bit measurements in the distributed setting.

\section{Adaptive Estimation \label{sec:sequential}}
The first main results of this paper, as described in Theorem~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be larger than $\eta(0)\sigma^2$, which is the ARE of the median of the sample $X_1,\ldots,X_n$. We also provide a particular adaptive estimation scheme that attains this maximal efficiency. %Finally, in Theorem~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the chosen message $B_i$  minimizes the MSE given $X_i$ and the previous $i-1$ messages. %Numerical simulations for the case of $f(x)$ the normal density function shows that the ARE of the estimator described by this scheme is $\eta(0)\sigma^2$.

\subsection{Limited efficiency in the adaptive setting}
Our first result asserts that the ARE of any adaptive encoding and estimation scheme is bounded from above by $\eta(0)\sigma^2$. %This result follows from the following theorem:
\begin{thm}[maximal relative effeciency] \label{thm:adpative_lower_bound}
Let ${\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Figure~\ref{fig:setup}-(ii). Assumes that $\pi(\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
n\mathbb E\left[ (\theta-{\theta}_n)^2 \right] \geq   \frac{n}{ 4f^2(0) n + I_0},
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

\subsubsection*{Proof}
%We first bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$. The result then follows using the van-Trees inequality which bounds from below the risk of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are in
See Appendix~\ref{proof:thm:adpative_lower_bound}.\par

Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
\[
n\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{1}{4f^2(0)}+O(n^{-1}),
\]
and 
\[
\ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
\]
It is possible to extend Theorem~\ref{thm:adpative_lower_bound} to any other loss functions using a more subtle version of the Van Tress inequality provided in \cite{efroimovich1980information}. See also \cite{DBLP:journals/corr/abs-1902-08582}. \par
%
Next, we provide an adaptive encoding and estimation scheme that attains the maximal ARE of $\eta(0)\sigma^2$.

\subsection{Asymptotically optimal estimator}
Let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive sequence. Consider the following estimator ${\theta}_n$ for $\theta$:  
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation}
B_n = B_n (X_n,\theta_{n-1}) =\sgn (X_n - \theta_{n-1}).
\end{equation}
Define the $n$th step estimation as
\begin{equation} \label{eq:sgd_est}
\bar{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n \right\}_{n\in \mathbb N}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
%
%
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with a finite location Fisher information, namely
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)}  \right)^2 f(x) dx < \infty.
\]
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
Then for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align} 
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
\\
& \qquad \qquad \qquad \qquad = \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right], \label{eq:attaining_LAM}
\end{align}
where $Z \sim \Ncal(0,1)$. 
%
%
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
 \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{n\eta(0)} + o(n^{-1}). 
\]
\end{enumerate}
\end{thm}

\subsubsection*{Proof}
See Appendix~\ref{proof:thm:sgd}. \par

Theorem~\ref{thm:sgd} implies that the estimator ${\theta}_n$ defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} attains the maximal ARE, as established by Theorem~\ref{thm:adpative_lower_bound}. %Furthermore, it is locally minimax in the sense that it attains the minimal MSE for all local alternatives of $\theta$. 
%
The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. In particular, parts (i) and (ii) of Theorem~\ref{thm:sgd} provides conditions under which $\bar{\theta}$ is asymptotically normal and local asymptotically minimax, in the sense that it attains the local asymptotic minimax bound \cite{van2000asymptotic}. Part (iii) provides an additional condition under which $\bar{\theta}$ also converges in its second moment. 
 \par
%Note that $\theta_0$ is not explicitly defined in equation \eqref{eq:sgd_est}. A reasonable initialization for $\theta_0$ is $\theta_0 = \mathbb E [\theta]$, although Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$.  \\
In the encoding and estimating procedure \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, each one-bit message $B_n$ depends on its private sample as well as the current gradient descent estimate $\theta_{n-1}$. In this sense, each encoder in this algorithm interacts with previous one by using the current estimate.
%
As we explain next, 
it is possible to obtain the optimal efficiency of $\eta(0)\sigma^2$ with only one round of interactions among the encoders.

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

% \node[below of = source, node distance = 1cm] (source2) {$X_2$};
%\node[int1, right of = source2, node distance = 1.2cm] (enc2) {Enc};  
%\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source, node distance = 1.7cm] (source3) {$X_{n_1}$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};

\draw[->,line width = 2pt] (source3) -- (enc3);

\node[below of = source, node distance = 0.5cm] {$\vdots$};

\node[int1, right of = enc3, node distance = 2.1cm ] (est) {$\est_1$};

\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);

%\draw[->,line width = 2pt] (enc2) -| node[above, xshift = -1cm] (mes2) {$B_2$} (est);

\draw[->,line width = 2pt] (enc3) -- node[above, xshift = 0cm]  {$B_{n_1}$} (est);

\node[below right = 0.75 and 1.5 of source3] (sourceB) {$X_{n_1 +1}$} ;
\node[int1, right of = sourceB, node distance = 1.7cm] (enc1B) {$\enc$};  
\draw[->,line width = 2pt] (sourceB) -- (enc1B);

\node[below of = sourceB, node distance = 1.7cm] (source3B) {$X_n$};
\node[int1, right of = source3B, node distance = 1.7cm] (enc3B) {$\enc$};  
\draw[->,line width = 2pt] (source3B) -- (enc3B); 
\node[below of = sourceB, node distance = 0.4cm] {$\vdots$};

\node[int1, right of = enc3B, node distance = 2.1cm ] (estB) {$\est_2$};

\draw[->,line width = 2pt] (enc1B) -| node[above, xshift = -0.5cm] {$B_{n_1+1}$} (estB);
\draw[->,line width = 2pt] (enc3B) -- node[above] {$B_n$} (estB);

\draw[->,line width = 1pt] (est.east) node[above, xshift  =0.5cm] {${\theta}_{n_1}$} -| (enc1B.north);

%\draw[->,line width = 1pt] (enc1B) -- (enc3B);

\draw[->,line width = 0.5pt] (est.east) -| +(1.3,-0.5) -- +(1.3,-2.5) -| (enc3B.north);

%\draw[->,line width = 1pt] (enc1B)  -- node[right] {${\theta}_{n_1}$} (enc3B);

\draw[->,line width = 0.5pt] (estB) -- +(0.8,0) node[right] {${\theta}_n$};
\node[below of = enc1B, node distance = 0.5cm] {$\vdots$};

\end{tikzpicture}
\end{center}
\caption{Distributed encoding with a single interaction: The estimation obtained from the first $n_1$ bits in a distributed manner is to obtain another $n-n_1$ bits in a distributed manner. 
\label{fig:one_round}
}
\end{figure}

\subsection{Maximal Efficiency using One Round of Threshold Adaptation}
In Section~\ref{sec:preliminary} we considered an estimator that is based on binary messages of the form 
\[
B_i = \mathbf 1_{X_i > \theta_0},\qquad i=1,\ldots,n,
\]
and deduced that it is asymptotically normal with variance $1/\eta(\theta-\theta_0)$.
%
We now show that a similar encoding leads to an asymptotically normal estimator attaining the lower variance bound $1/\eta(0)$, provided we allow to  update the threshold value $\theta_0$ based on previously observed bits at least once. 
%
In this procedure we separate the sample into two disjoint sets: $X_1,\ldots,X_{n_1}$ and $X_{n_1+1},\ldots,X_n$ for some $n_1 < n$.
We first use the estimator \eqref{eq:estimator_naive} to obtain an estimate ${\theta}_{n_1}$ based on $B_1,\ldots,B_{n_1}$, and then use ${\theta}_{n_1}$ as the new threshold value to obtain messages $B_{n_1+1}, \ldots, B_n$. Figure~\ref{fig:one_round} illustrates a diagram of this procedure. 
%
The specific encoding and estimation scheme, as well as its asymptotic performance, are given by the following theorem:
\begin{thm}
For $i=1,\ldots,n$ set
\[
B_i = \begin{cases}
 \mathbf 1_{X_i \geq \theta_0} & i = 1,\ldots,n_1, \\
 \mathbf 1_{X_i \geq {\theta}_{n_1} }& i={n_1+1,\ldots,n},
\end{cases}
\]
where $n = n_1+ n_2$ and 
\[
{\theta}_{n_1} \triangleq \theta_0 + F^{-1}\left(
\frac{1}{n_1} \sum_{i=1}^{n_1} B_i 
 \right).
\] 
Let 
\[
{\theta}_{n_2}  = {\theta}_{n_1} +  F^{-1} \left( \frac{1}{n_2} \sum_{i=n_1+1}^{n_2} B_i \right)
\]
and assume that $n_1(n) \rightarrow \infty$ and $n_1(n)/n \rightarrow 0$. Then:
\begin{align*}
 \sqrt{n} \left( {\theta}_{n_2} - \theta  \right) \overset{d}{\longrightarrow}  \Ncal\left( 0, 1/\eta(0) \right).
\end{align*}
\end{thm}
%
%
\begin{proof}
For $t\in \mathbb R$, set
\[
p_{n_2}(t) \triangleq \frac{1}{n_2} \sum_{i=n_1+1}^{n} \mathbf 1_{X_i \geq t}. 
\]
From the central limit theorem
\begin{align*}
& \sqrt{n} \left( p_{n_2}(t) - F( \theta - t) \right)  = \sqrt{\frac{n}{n_2}} \sqrt{n_2} \left(  p_{n_2}(t) - F(\theta - t) \right) \\
& \overset{d}{\rightarrow} \Ncal\left( 0, V(t) \right),
\end{align*}
where
\[
V(t) \triangleq F(\theta - t) F \left( t - \theta\right). 
\]
Applying the delta method to $p_{n_2}(t)$ with $g(x) = F^{-1}(x)$ implies
\begin{align*}
& \sqrt{n} \left( t + F^{-1}(\hat{p}_{n_2}(t)) - \theta \right) \\
&= \sqrt{n} \left( g(\hat{p}_{n_2}(t)) - g \left(F(\theta-t) \right)  \right) \\
& \overset{d}{\to} \Ncal\left(0, 
1/\eta(\theta-t) \right).
\end{align*}
%where this convergence is uniform in $t$ since all moments of $F^{-1}(\hat{p}_{n_2})$ exists. 
On the other hand, by the law of large numbers,
\[
p_{n_1} \triangleq \frac{1}{n_1} \sum_{i=1}^{n_1} B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),
\]
so that ${\theta}_{n_1}$ converges almost surely to $\theta$ as $n$ goes to infinity and thus 
\[
\eta( {\theta}_{n_1}-\theta) \overset{a.s.}{\rightarrow} \eta(0). 
\]
Finally,
\begin{align*}
& \sqrt{n}\left({\theta}_{n_2} - \theta\right) \\
& = \sqrt{n} \left( {\theta}_{n_1} + F^{-1}(\hat{p}_{n_2}({\theta}_{n_1})) - \theta \right)  \\
& \overset{d}{\to} \Ncal\left( 0, 
1/\eta(0) \right) 
\end{align*}
by Slutsky's theorem.
\end{proof}
%

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=10cm, height=6cm,
xmin = 0, xmax=800, 
restrict y to domain = 0:3,
ymin = 0,
ymax = 3.4,
samples=10, 
xlabel= $n$,
ylabel = {$n\mathbb E \left[\left(\theta - {\theta}_n \right)^2 \right]$},
%xtick={-3,-2,-1,0,1,2,3},
%xticklabels={-3,-2,-1,0,1,2,3},
ytick={0,1,1.57},
yticklabels={0,1,$\pi/2$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]

\addplot[color = blue, solid, smooth] plot table [x = itr, y = SGD, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
\addlegendentry{asymptotically optimal};

>>>>

% -*- Mode: latex -*- %
